### 2024-08-11 doc/presentation.qmd
rendering presentation

---
title: |
  | Earnings Management and Investor Protection:
  | Accounting Reading Group - Assignment III
author: "Melisa Mazaeva"
institute: "Humboldt-Universität zu Berlin – Summer 2024"
date: today
fontsize: "9pt"
pdf-engine: xelatex # pdflatex creates rastered fonts
format: 
  beamer:
    slide-level: 3
    number-sections: true
    toc: false
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\fontsize{7}{7}\selectfont}  # This sets the font size for verbatim environments
- \input{beamer_theme_trr266.sty}
bibliography: references.bib
biblio-style: apsr
---

```{python}
#| echo: false
#| output: false

import pickle

with open('../output/results.pickle', 'rb') as f:
        results = pickle.load(f)

min_fyear, max_fyear, no_unique_firms = results['Desc information'].values()

def rename_var_to_label(table, italic=True):
    for var_name, var_label in results['Variable Names'].items():
        var_label = f'\\textit{{{var_label}}}' if italic else var_label
        table = table.replace(var_name, var_label)
    return table
```

\pagebreak

# Motivation

### Discretionary Accruals

- Are used in tons of papers
- As measurement constructs for almost everything
  - Earnings Management
  - Earnings Quality
  - Audit Quality
  - Management Quality
- But how do they look like?

### Data

```{python}
#| echo: false
#| output: asis

lat_tab = rename_var_to_label(results['Descriptive Statistics']).split('\n')
# lat_tab[4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
# lat_tab[6] = '\\hline\\\\[-1.8ex]'
# lat_tab[-4] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'

print('\n'.join([
    '\\begin{table}[!htbp]',  
    '\\resizebox{\\textheight}{!}{',
    '\\begin{threeparttable}',
    '\n'.join(lat_tab[3:-2]),
    '\\begin{tablenotes}[flushleft]',
    '\setlength{\labelsep}{0pt}',
    f'\\item Note: The data is obtained from the Compustat U.S. as provided by WRDS. The sample covers the period {int(min_fyear)} to {int(max_fyear)} and {no_unique_firms:,} unique firms.',
    '\end{tablenotes}',
    '\end{threeparttable}',
    '}',
    '\\end{table}',
]))
```

### Distribution of Discretionary Accruals over time - unrestricted sample

```{python}
#| label: fig-boxplot_full
#| fig-cap: "Distribution of Discretionary Accruals over time - unrestricted sample"
#| echo: false
results['fig_boxplot_full'].set_figheight(2.6)
results['fig_boxplot_full']
```

### Distribution of Discretionary Accruals over time - analysis sample

```{python}
#| label: fig-boxplot_smp
#| fig-cap: "Distribution of Discretionary Accruals over Time - analysis sample"
#| echo: false
results['fig_boxplot_smp'].set_figheight(2.6)
results['fig_boxplot_smp']
```

### Correlation of Modified Jones with Dechow and Dichev DA

```{python}
#| label: fig-scatter_md_dd
#| fig-cap: "Correlation of Modified Jones with Dechow and Dichev DA"
#| echo: false
results['fig_scatter_md_dd'].set_figheight(3)
results['fig_scatter_md_dd']
```

### Association of Dechow and Dichev DA with firm size

```{python}
#| label: fig-scatter_dd_lnta
#| fig-cap: "Dechow and Dichev DA and Firm Size"
#| echo: false

results['fig_scatter_dd_lnta'].set_figheight(3)
results['fig_scatter_dd_lnta']
```

### No presentation without correlation table...

```{python}
#| echo: false
#| output: asis

lat_tab = rename_var_to_label(results['Correlations'], False).split('\n')
lat_tab[2] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'
lat_tab[4] = '\\hline\\\\[-1.8ex]'
lat_tab[-7] = '\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]'

print('\n'.join([
    '\\begin{center}',
    '\\resizebox*{\\textwidth}{!}{',
    '\\begin{tabular}{lccccccc}',
    '\n'.join(lat_tab[2:-6]),
    '\\end{tabular}}',
    '\\end{center}'
]))
```

### ... and regression table

```{python}
#| echo: false
#| output: asis

print('\n'.join([
    '\\begin{center}',
    '\\resizebox*{!}{0.8\\textheight}{', 
    '\\begin{tabular}{lcc}',
    rename_var_to_label('\n'.join(results['Regression'].split('\n')[4:-2])),
    '\\end{tabular}}',
    '\\end{center}'
]))
```

### Conclusion

- Discretionary accruals are simply wonderful! 
- With these distributional properties and the power of large samples they will almost always be significant in one direction or the other regardless the setting or controls
- Use them for every economic construct under the sun!